### 2010-10-19 papers/BMD2010/PhysicalParametersPaper/paper.tex
editing the document

\begin{center}
\begin{tabular}{c}
$^*$ Department of Mechanical and Aerospace Engineering\\
University of California, Davis\\
One Shields Avenue, Davis, CA, 95616, USA\\
e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu\\
\end{tabular}
\begin{tabular}{c}
$^\dag$ Laboratory for Engineering Mechanics\\
Delft University of Technology\\
Mekelweg 2, 2628CD Delft, The Netherlands\\
e-mail: a.l.schwab@tudelft.nl, jodikooijman@gmail.com\\
\end{tabular} \\ \vspace{2.5ex}
\end{center}
\section*{ABSTRACT}
Accurate measurements of a bicycle's physical parameters are required for
realistic dynamic simulations and analysis. The most basic models require the
geometry, mass, mass location and mass distributions for the rigid bodies. More
complex models require estimates of tire characteristics, human
characteristics, friction, stiffness, damping, etc. In this
paper we present the measurement of the minimal bicycle parameters required for
the benchmark Whipple bicycle model presented in~\cite{Meijaard2007}. This
model is composed of four rigid bodies, has ideal rolling and frictionless joints,
and is laterally symmetric. A set of 25 parameters is used to describe the
geometry, mass, mass location and mass distribution of each of the rigid
bodies. The experimental methods used to estimate the parameters described
herein are based primarily on the work done in~\cite{Kooijman2006} but have
been refined for improved accuracy and methodology.

Koojiman's work was preceded by \cite{Roland1971} who measured a bicycle in a
similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
similar techniques with scooters. We measured the physical characteristics of six
different bicycles, two of which were set up in two different configurations.

The six bicycles, chosen for both variety and convenience, are as follows:
\emph{Batavus Browser}, a Dutch style city bicycle measured with and without
instrumentation as described in~\cite{Kooijman2009a}; \emph{Batavus Stratos
Deluxe}, a Dutch style sporty city bicycle; \emph{Batavus Crescendo Deluxe} a
Dutch style city bicycle with a suspended fork; \emph{Gary Fisher Mountain
Bike}, a hardtail mountain bicycle; \emph{Bianchi Pista}, a modern steel frame
track racing bicycle; and \emph{Yellow Bicycle}, a stripped down aluminum frame
road bicycle measured in two configurations, the second with the fork rotated
in the headtube 180 degrees for larger trail.

These eight different parameter sets can be used with, but not
limited to, the benchmark bicycle model. The accuracy of all the measurements
are presented up through the eigenvalue prediction of the linear model. The
accuracies are based on error propagation theory with correlations taken into account.
\begin{keywords}
bicycle,
parameters,
eigenvalues,
Bode.
\end{keywords}

\section{INTRODUCTION}

This work is intended to document the indirect measurement of eight real
bicycles' physical parameters. The physical parameters measured are those that
are needed for the benchmark bicycle model~\cite{Meijaard2007}. The work is
based off of the techiques used to measure the instrumented bicycle in
\cite{Kooijman2006} and \cite{Kooijman2008}. I used the same techniques as
Kooijman to measure a second instrumented bicycle~\cite{Moore2009a}. This
method had low accuracy and the $I_{yy}$ moments of inertia were
not measured for the fork and frame. Furthermore, very little data exists on
the physical parameters of different types of bicycles and this work aims to
provide a small sample of bicycles.

>>>>

\begin{center}
\begin{tabular}{c}
$^*$ Department of Mechanical and Aerospace Engineering\\
University of California, Davis\\
One Shields Avenue, Davis, CA, 95616, USA\\
e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu\\
\end{tabular}
\begin{tabular}{c}
$^\dag$ Laboratory for Engineering Mechanics\\
Delft University of Technology\\
Mekelweg 2, 2628CD Delft, The Netherlands\\
e-mail: a.l.schwab@tudelft.nl, jodikooijman@gmail.com\\
\end{tabular} \\ \vspace{2.5ex}
\end{center}
\section*{ABSTRACT}
Accurate measurements of a bicycle's physical parameters are required for
realistic dynamic simulations and analysis. The most basic models require the
geometry, mass, mass location and mass distributions for the rigid bodies. More
complex models require estimates of tire characteristics, human
characteristics, friction, stiffness, damping, etc. In this
paper we present the measurement of the minimal bicycle parameters required for
the benchmark Whipple bicycle model presented in~\cite{Meijaard2007}. This
model is composed of four rigid bodies, has ideal rolling and frictionless joints,
and is laterally symmetric. A set of 25 parameters is used to describe the
geometry, mass, mass location and mass distribution of each of the rigid
bodies. The experimental methods used to estimate the parameters described
herein are based primarily on the work done in~\cite{Kooijman2006} but have
been refined for improved accuracy and methodology. Koojiman's work was
preceded by \cite{Roland1971} who measured a bicycle in a
similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
similar techniques with scooters.

We measured the physical characteristics of six different bicycles, two of
which were set up in two different configurations. The six bicycles, chosen for
both variety and convenience, are as follows:
\emph{Batavus Browser}, a Dutch style city bicycle measured with and without
instrumentation as described in~\cite{Kooijman2009a}; \emph{Batavus Stratos
Deluxe}, a Dutch style sporty city bicycle; \emph{Batavus Crescendo Deluxe} a
Dutch style city bicycle with a suspended fork; \emph{Gary Fisher Mountain
Bike}, a hardtail mountain bicycle; \emph{Bianchi Pista}, a modern steel frame
track racing bicycle; and \emph{Yellow Bicycle}, a stripped down aluminum frame
road bicycle measured in two configurations, the second with the fork rotated
in the headtube 180 degrees for larger trail.

These eight different parameter sets can be used with, but not
limited to, the benchmark bicycle model. The accuracy of all the measurements
are presented up through the eigenvalue prediction of the linear model. The
accuracies are based on error propagation theory with correlations taken into
account.

\begin{keywords}
bicycle,
parameters,
eigenvalues,
Bode.
\end{keywords}

\section{INTRODUCTION}
This work is intended to document the indirect measurement of eight real
bicycles' physical parameters. The physical parameters measured are those that
are needed for the benchmark Whipple bicycle model presented in
\cite{Meijaard2007}. The work is based on the techiques used to measure the
instrumented bicycles in \cite{Kooijman2006}, \cite{Kooijman2008} and
\cite{Moore2009a}. We improve upon these methods by both increasing and reporting
the accuracies of the measurements and by measuring the complete moments of
inertia of the frame and fork needed for anaylsis of the nonlinear model.
Furthermore, very little data exists on the physical parameters of different
types of bicycles and this work aims to provide a small sample of bicycles.
